---
title: "Create data for rating scale analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Overview

This script uses the `psych` package to generate two 1000-case data sets for  demonstrating the procedures of rating scale analysis. The first, `data_input_sim`, is simulated data consisting of 50 behavior items rated on a `Never-Occasionally-Frequently-Always` scale. These responses are coded `1-2-3-4` and summed to yield five raw scale scores: `S1_raw`, `S2_raw`, `S3_raw`, `S4_raw`, and `S5_raw`. The data set also includes simulated data for age, parental education level, ethnicity, geographic region, and clinical status. 

The second data set, `data_input_bfi`, is real data consisting of 25 self-report items taken from the International Personality Item Pool. These responses are coded `1-2-3-4-5-6` for level of agreement with the personality descriptor. The item codes are summed to yield five raw scale scores: `AGR_raw`, `CON_raw` `EXT_raw`, `NEU_raw`, and `OPE_raw`.  This data set includes actual data for age, gender, and education; and simulated data for ethnicity, region, and clinical status.

#### Executable Code
```{r create-data-rsa, eval=FALSE}
suppressMessages(library(here))
suppressMessages(library(tidyverse))
suppressMessages(library(psych))

set.seed(123)

data_input_sim <-
  as_tibble(sim.poly.ideal(nvar = 50, n = 1000, cat = 4,)[["items"]]) %>%
  rename_with(~ str_c("i", str_pad(
    as.character(1:50), 2, side = "left", pad = "0"
  ))) %>%
  mutate(
    S1_raw = rowSums(.[str_c("i", str_pad(as.character(seq(1, 50, by = 5)), 2, side = "left", pad = "0"))]),
    S2_raw = rowSums(.[str_c("i", str_pad(as.character(seq(2, 50, by = 5)), 2, side = "left", pad = "0"))]),
    S3_raw = rowSums(.[str_c("i", str_pad(as.character(seq(3, 50, by = 5)), 2, side = "left", pad = "0"))]),
    S4_raw = rowSums(.[str_c("i", str_pad(as.character(seq(4, 50, by = 5)), 2, side = "left", pad = "0"))]),
    S5_raw = rowSums(.[str_c("i", str_pad(as.character(seq(5, 50, by = 5)), 2, side = "left", pad = "0"))]),
    across(
      contains("i"),
      ~ case_when(
        .x == 0 ~ "never",
        .x == 1 ~ "occasionally",
        .x == 2 ~ "frequently",
        .x == 3 ~ "always"
      )
    ),
    ID = 100001:101000,
    age = sample(c(5:12), 1000, replace = TRUE),
    age_range = case_when(age <= 8 ~ "5 to 8 yo",
                          T ~ "9 to 12 yo"),
    gender = sample(
      c("female", "male"),
      1000,
      replace = TRUE,
      prob = c(0.53, 0.47)
    ),
    educ = sample(
      c("no_HS", "HS_grad", "some_college", "BA_plus"),
      1000,
      replace = TRUE,
      prob = c(0.119, 0.263, 0.306, 0.311)
    ),
    ethnic = sample(
      c("hispanic", "asian", "black", "white", "other"),
      1000,
      replace = TRUE,
      prob = c(0.239, 0.048, 0.136, 0.521, .056)
    ),
    region = sample(
      c("northeast", "south", "midwest", "west"),
      1000,
      replace = TRUE,
      prob = c(0.166, 0.383, 0.212, 0.238)
    ),
    clin_status = sample(
      c("typ", "clin"),
      1000,
      replace = TRUE,
      prob = c(0.8, 0.2)
    )
  ) %>%
  select(ID:clin_status, S1_raw:S5_raw, i01:i50)

write_csv(data_input_sim,
          here("INPUT-FILES/data-input-sim.csv"),
          na = "")

data_input_bfi <- bfi %>%
  drop_na() %>%
  sample_n(1000) %>%
  mutate(
    AGR_raw = rowSums(.[str_c("A", 1:5)]),
    CON_raw = rowSums(.[str_c("C", 1:5)]),
    EXT_raw = rowSums(.[str_c("E", 1:5)]),
    NEU_raw = rowSums(.[str_c("N", 1:5)]),
    OPE_raw = rowSums(.[str_c("O", 1:5)]),
    across(
      c(A1:O5),
      ~
        case_when(
          .x == 1 ~ "very_inaccurate",
          .x == 2 ~ "moderately_inaccurate",
          .x == 3 ~ "slightly_inaccurate",
          .x == 4 ~ "slightly_accurate",
          .x == 5 ~ "moderately_accurate",
          .x == 6 ~ "very_accurate",
        )
    ),
    ID = 200001:201000,
    age_range = case_when(
      age <= 18 ~ "18 yo or younger",
      between(age, 19, 24) ~ "19 to 24 yo",
      between(age, 25, 39) ~ "25 to 39 yo",
      T ~ "40 yo or older"
    ),
    gender = case_when(gender == 1 ~ "male",
                       gender == 2 ~ "female"),
    educ = case_when(
      education == 1 ~ "no_HS",
      education == 2 ~ "HS_grad",
      education == 3 ~ "some_college",
      T ~ "BA_plus"
    ),
    ethnic = sample(
      c("hispanic", "asian", "black", "white", "other"),
      1000,
      replace = TRUE,
      prob = c(0.239, 0.048, 0.136, 0.521, .056)
    ),
    region = sample(
      c("northeast", "south", "midwest", "west"),
      1000,
      replace = TRUE,
      prob = c(0.166, 0.383, 0.212, 0.238)
    ),
    clin_status = sample(
      c("typ", "clin"),
      1000,
      replace = TRUE,
      prob = c(0.8, 0.2)
    )
  ) %>%
  select(ID, age, age_range, gender:clin_status, AGR_raw:OPE_raw, A1:O5)

write_csv(data_input_bfi,
          here("INPUT-FILES/data-input-bfi.csv"),
          na = "")

rm(data_input_bfi)
```

<br>

#### Commented
