---
title: "Rating Scale Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### t-score comparisons between clinical group, matched typical group

A core set of analyses supporting rating scale validity is the comparison of clinical groups (e.g., those with diagnoses of ASD, ADHD, LD, etc.) to groups of typically developing students who are matched on demographic characteristics. For this demonstration, we employ data collected in ???? as part of the SPM-2 project. We have a clinical group consisting of children who have been diagnosed with ASD. The goal of the validity analysis is to determine the extent to which the SPM-2 t-scores of this clincial group differ from a group of typically developing children with matched demographic characteristics.

The method therefore focuses on drawing a matched subsample from a larger sample of typically developing children (in this case the SPM-2 Standardization Sample). Once we have assemble comparison group, we then calculate descriptive statistics on each group, and use a measure of effect size to assess the magnitude of differences between the two groups on SPM-2 scores.

Output includes two tables:

-   Demographic characteristics
-   Descriptive statistics with effects sizes for the means comparisons

This demonstration that you have set up an RStudio project called `RATING-SCALE-ANALYSIS`, with subfolders `CODE`, `INPUT-FILES` and `OUTPUT-FILES`. Save scripts in `CODE`, and run them from within the `RATING-SCALE-ANALYSIS` project.

###### EXECUTABLE CODE

```{r RSA-match, eval = FALSE}
suppressMessages(library(here)) 
suppressMessages(suppressWarnings(library(tidyverse)))
suppressMessages(library(psych))
suppressMessages(library(MatchIt))

var_order <- c("age_range", "Age", "AgeInMonths", "Gender", "ParentHighestEducation", "HighestEducation", 
               "Ethnicity", "Region")
cat_order <- c(
  # AgeInMonths
  NA, '1', '2', '3', '4', '5', '6', '7', '8', '9', '10', '11', '12', '13', '14', '15', '16', '17', '18', '19', '20', '21', '22', 
  '23', '24', '25', '26', '27', '28', '29', '30', '31', '32', '33', '34', '35', '36',
  # age_range
  NA, "03.5 to 6 mo", "03.5 to 10 mo", "4 to 10 mo", "07 to 10.5 mo", "09.5 to 20 mo",  
  "11 to 20 mo", "11 to 31.5 mo", "21 to 30 mo",
  "21 to 31.5 mo", "2 to 4 years", "5 years", "5 to 8 years", "9 to 12 years", "12 to 13 years", 
  "14 to 15 years", "16 to 17 years", "18 to 21 years", "21.00 to 30.99 years", "31.00 to 40.99 years", 
  "41.00 to 50.99 years", "51.00 to 64.99 years", "65.00 to 99.99 years",
  # Age
  "2", "3", "4", "5",
  # Gender
  NA, "Male", "Female",
  # ParentHighestEducation & HighestEducation
  NA, "Did not complete high school (no diploma)", "High school graduate (including GED)", 
  "Some college or associate degree", "Bachelor's degree or higher",
  # Ethnicity
  NA, "Hispanic", "Asian", "Black", "White", "AmericanIndAlaskanNat", 
  "NativeHawPacIsl", "MultiRacial", "Other",
  # Region
  NA, "northeast", "midwest", "south", "west")

stand_preMatch <- 
  suppressMessages(as_tibble(read_csv(
    here("INPUT-FILES/data-RS-spm2-stand-all-T-scores-per-case.csv")
  )))

clin_ASD_preMatch <-
  suppressMessages(as_tibble(read_csv(
    here("INPUT-FILES/data-RS-spm2-clin-all-T-scores-per-case.csv")
  ))) %>% 
  filter(clin_dx == "ASD")

ASD_clin_stand_preMatch <- bind_rows(
  clin_ASD_preMatch,
  stand_preMatch 
) %>% 
  mutate(Group = case_when(
    clin_status == 'clin' ~ TRUE,
    TRUE ~ FALSE
  ))

sum(is.na(ASD_clin_stand_preMatch))
na_cols <- ASD_clin_stand_preMatch %>% select_if(~ any(is.na(.)))

ASD_clin_stand_preMatch <- ASD_clin_stand_preMatch %>%
  replace_na(
    list(
      IDNumber = 999,
      AgeInMonths = 999,
      Age = 999,
      clin_dx = "999",
      Region = "999"
    )) %>% 
  mutate(HighestEducation = replace_na(as.character(HighestEducation), "999"))

set.seed(12345)
match <- matchit(
  Group ~ age_range + Gender + ParentHighestEducation + Ethnicity, 
  data = ASD_clin_stand_preMatch, 
  method = "nearest", 
  ratio = 1)
match_summ <- summary(match)

ASD_clin_stand_match <- match.data(match) %>% 
  select(-Group, -distance, -weights) #%>% 
ASD_stand_match <- ASD_clin_stand_match %>% 
  filter(clin_status == 'typ')
ASD_clin_match <- ASD_clin_stand_match %>% 
  filter(clin_status == 'clin')

match_dist_stand <- ASD_stand_match %>% 
  select(age_range, Gender, ParentHighestEducation, Ethnicity) %>% 
  gather("Variable", "Category") %>% 
  group_by(Variable, Category) %>%
  count(Variable, Category) %>%
  arrange(match(Variable, var_order), match(Category, cat_order)) %>% 
  ungroup() %>% 
  mutate(Variable = case_when(
    lag(Variable) == "age_range" & Variable == "age_range" ~ "",
    lag(Variable) == "Gender" & Variable == "Gender" ~ "",
    lag(Variable) == "ParentHighestEducation" & Variable == "ParentHighestEducation" ~ "",
    lag(Variable) == "Ethnicity" & Variable == "Ethnicity" ~ "",
    TRUE ~ Variable
  )) %>% 
  mutate(group = case_when(
    rownames(.) == "1" ~ 'Matched typical',
    T ~ NA_character_
  )) %>% 
  select(group, everything())

match_dist_clin <- ASD_clin_match %>% 
  select(age_range, Gender, ParentHighestEducation, Ethnicity) %>% 
  gather("Variable", "Category") %>% 
  group_by(Variable, Category) %>%
  count(Variable, Category) %>%
  arrange(match(Variable, var_order), match(Category, cat_order)) %>% 
  ungroup() %>% 
  mutate(Variable = case_when(
    lag(Variable) == "age_range" & Variable == "age_range" ~ "",
    lag(Variable) == "Gender" & Variable == "Gender" ~ "",
    lag(Variable) == "ParentHighestEducation" & Variable == "ParentHighestEducation" ~ "",
    lag(Variable) == "Ethnicity" & Variable == "Ethnicity" ~ "",
    TRUE ~ Variable
  )) %>% 
  mutate(group = case_when(
    rownames(.) == "1" ~ 'ASD',
    T ~ NA_character_
  )) %>% 
  select(group, everything())

match_dist <- bind_rows(
  match_dist_clin,
  match_dist_stand
)

ASD_matchStand <- stand_preMatch %>% 
  semi_join(ASD_stand_match, by ='IDNumber')

scale_order <- c("SOC", "VIS", "HEA", "TOU", 
                 "TS", "BOD", "BAL", "PLA", "TOT")

ASD_matchStand_t_desc <-
  ASD_matchStand %>% 
  select(contains('_NT')) %>% 
  rename_with(~ str_sub(., 1, -4), everything()) %>% 
  describe(fast = T) %>%
  rownames_to_column() %>% 
  rename(scale = rowname)%>% 
  arrange(match(scale, scale_order)) %>% 
  mutate(sample = case_when(
    rownames(.) == "1" ~ 'Matched typical',
    T ~ NA_character_
  ),
  var = sd^2) %>% 
  select(sample, scale, n, mean, sd, var) %>% 
  rename(n_typ = n,
         mean_typ = mean,
         sd_typ = sd,
         var_typ = var)

ASD_clin_t_desc <-
  clin_ASD_preMatch %>% 
  select(contains('_NT')) %>% 
  rename_with(~ str_sub(., 1, -4), everything()) %>% 
  describe(fast = T) %>%
  rownames_to_column() %>% 
  rename(scale = rowname)%>% 
  arrange(match(scale, scale_order)) %>% 
  mutate(sample = case_when(
    rownames(.) == "1" ~ 'ASD',
    T ~ NA_character_
  ),
  var = sd^2) %>% 
  select(sample, scale, n, mean, sd, var) %>% 
  rename(n_clin = n,
         mean_clin = mean,
         sd_clin = sd, 
         var_clin = var)

ASD_match_t_desc <- left_join(ASD_matchStand_t_desc,
                              ASD_clin_t_desc, by = "scale") %>%
  mutate(ES = abs((mean_typ - mean_clin) / sqrt(((n_typ * var_typ) + (n_clin * var_clin)) / (n_typ + n_clin))),
         across(c(mean_typ, sd_typ, mean_clin, sd_clin, ES), ~
                  (round(., 2)))) %>%
  select(everything(), -sample.x, -sample.y, -var_typ, -var_clin)

write_csv(match_dist,
          here(
            "OUTPUT-FILES/TABLES/demos-ASD-matchTyp.csv"
          ),
          na = '')

write_csv(ASD_match_t_desc,
          here(
            "OUTPUT-FILES/TABLES/t-desc-ES-ASD-matchTyp.csv"
          ),
          na = '')
```

<br>

###### COMMENTED SNIPPETS


#### 1. Load packages, initialize tokens for robust code, read data

Load packages for file path specification (`here`), data wrangling (`tidyverse`), psychometric data simulation and analysis (`psych`), and drawing matched samples (`MatchIt`).

Ideally, code templates are *robust*, meaning that they can be adapted to different projects with minimal copying-and-pasting of text. Robust templates are characterized by their liberal use of *tokens*, which are names for *project-specific* data and text elements that are used repeatedly throughout the template.

These tokenized elements include names of variables, files, test forms, raters, etc., as well as labels for age- and score ranges, score item counts, score numerical bounds, and so forth. By definition, these elements differ between, say, an autism rating-scale project and an ADHD rating-scale project.

In contrast, the *project-general* aspects of an R script are those functions and operations whose specification is identical for all projects (e.g., creating a table of descriptive statistics requires calling the same set of functions, regardless of whether the input data are from the autism rating scale or the ADHD rating scale). In a robust template, tokens representing project-specific elements are defined at the head of the script. The remainder of the script, therefore, consists only of project-general code.

In R, we initialize tokens as *vectors*. In the next code snippet, for example, `var_order` is a vector specifying the sequence in which the names of demographic variables are presented in the output tables.

In this demonstration, we employ data from the SPM-2 project, because that project has both small clinical samples and a large standardization sample from which we can draw matched subsamples. We read in the data using `read_csv()`, transforming the `.csv` files into the simplified tibble format with `as_tibble().` The data are in two objects named `stand_preMatch` and `clin_ASD_preMatch`. The names differentiate the clinical group (children diagnosed with ASD) from the typically developing standardization group, and indicate that the data are in their raw input form, prior to the matched sample draw (`preMatch`). For the clinical group, we use
`filter(clin_dx == "ASD")` to read in only the persons with ASD, out of the larger clinical sample that contains numerous diagoses.

```{r RSA-match, echo = 1:40, eval = F}
```

#### 2. Prep input data for matching

Text
```{r RSA-match, echo = 42:63, eval = F}
```


