---
title: "Rating Scale Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Check for duplicate, missing IDs

This code processes a single input file and generates a .csv report that includes rows with missing or duplicated case IDs. 

###### EXECUTABLE CODE
```{r dup, eval = FALSE}
suppressMessages(library(here))
suppressMessages(suppressWarnings(library(tidyverse)))

urlRemote_path  <- "https://raw.github.com/"
github_path <- "wpspublish/DSHerzberg-RATING-SCALE-ANALYSIS/master/INPUT-FILES/"

data_RS_sim_child_parent <- suppressMessages(read_csv(url(
  str_c(urlRemote_path, github_path, "data-RS-sim-child-parent.csv")
)))

miss_dup <- data_RS_sim_child_parent %>%
  mutate(across(ID,
                ~
                  case_when(
                    row_number() %in% c(10, 35, 579, 991) ~ lag(.x),
                    row_number() %in% c(11) ~ lag(lag(.x)),
                    row_number() %in% c(1, 2, 414, 793) ~ NA_real_,
                    T ~ .x
                  )))

anyDuplicated(miss_dup$ID)
any(is.na(miss_dup$ID))

data_RS_sim_child_parent_dupMissIDs <- miss_dup %>%
  arrange(ID) %>% 
  mutate(dup = duplicated(ID)) %>%
  filter(dup == T | lead(dup) == T | is.na(ID)) %>%
  select(-dup)

write_csv(data_RS_sim_child_parent_dupMissIDs,
  here(
    str_c(
      'OUTPUT-FILES/DUP-IDS/data-RS-sim-child-parent-dupIDs-missingIDs-',
      format(Sys.Date(), "%Y-%m-%d"),
      '.csv'
    )
  ), 
  na = 'missing'
)
```

###### COMMENTED SNIPPETS
Load packages for file path specification (`here`) and data wrangling (`tidyverse`).
```{r dup, echo = 1:2, eval = F}
```
Specify file paths and retrieve data from a remote host. 
```{r dup, echo = 4:9, eval = F}
```
Recode `ID` to include missing and duplicate values, for demonstration purposes, and hold the recoded data frame in `miss_dup`. In this example, one duplicate ID appears three times. This code chunk should be commented out when processing actual data.
```{r dup, echo = 11:19, eval = F}
```
Check for the presence of duplicate and/or missing case IDs.

* `base::anyDuplicated()`
  + If duplicate case IDs are present, returns the row number of the first duplicate.
  + If no duplicates, returns `0`.
* `base::any(base::is.na())`
  + If at least one case ID is missing, returns `TRUE`.
  + If no missing case IDs, returns `FALSE`.
  
If `ID` has no duplicate or missing values, the analysis stops here.
```{r dup, echo = 21:22, eval = F}
```
Prepare a report containing rows with duplicate and/or missing case IDs. Sorting the input `miss_dup` by `ID` (with `dplyr::arrange()`) places any rows missing on `ID` at the tail of the data frame, ensuring that rows immediately preceding those that are `NA` on `ID` will not be mislabeled as duplicates.

`dplyr::mutate()` creates a `dup` column. It populates that column with `base::duplicated(ID)`, which returns `TRUE` for any row containing an additional instance of an `ID` value (inluding `NA`) that appeared in a previous row.

`dplyr:filter()` keeps only rows needed for the output report, i.e., those where:

* `dup` is `TRUE`
* `dup` is `TRUE` in the _next_ row
* `ID` is `NA`

The second logical condition is required because `duplicated(ID)` returns `FALSE` for the FIRST instance of a duplicated case ID. Because we sorted the input data frame by `ID`, however, duplicate values of `ID` appear in adjacent rows, so the first instance of a duplicate is captured when the predicate `lead(dup) == T` returns `TRUE`.
```{r dup, echo = 24:29, eval = F}
```
The remainder of the code writes the output report to .csv, using `readr::write_csv()`. Some additional detail:

* `format(Sys.Date(), "%Y-%m-%d")` returns the current data in ISO 8601 format. This date is appended to the output file name with `str_c()`.
* `na = ""`: writes `NA` values as blank cells in the output .csv.

