---
title: "Rating Scale Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Create Data

#### Overview

We use the `psych` package to generate two 1000-case data sets for  demonstrating the procedures of rating scale analysis. The first, `data_RS_sim_child_parent`, is simulated data consisting of 50 behavior items rated on a four-option `Never-Occasionally-Frequently-Always` scale. The data are parent ratings on children aged 5-12. The item and scale score names have a `cp` (upper or lower case) prefix to indicate `parent` ratings on a `child` form. By varying these prefixes, and changing the definitions of the `age` and `age_range` variables, we can create similar data sets for other forms (e.g., `teacher`) and other age ranges (e.g., `teen`).

The item responses are coded `1-2-3-4` and summed to yield a raw total score `CPTOT_raw`, and five raw subscale scores: `CPS1_raw`, `CPS2_raw`, `CPS3_raw`, `CPS4_raw`, and `CPS5_raw`. The data set also includes simulated data for age, parental education level, ethnicity, geographic region, and clinical status. 

The second data set, `data_input_bfi`, is real data consisting of 25 self-report items taken from the International Personality Item Pool. These responses are coded `1-2-3-4-5-6` for level of accuracy of the personality descriptor, with higher codes indicating greater accuracy. The item codes are summed to yield five raw scale scores: `AGR_raw`, `CON_raw` `EXT_raw`, `NEU_raw`, and `OPE_raw`.  This data set includes real data for age, gender, and education; and simulated data for ethnicity, region, and clinical status.

#### Executable Code
```{r create-data-rsa, eval = FALSE}
suppressMessages(library(here))
suppressMessages(library(tidyverse))
suppressMessages(library(psych))

set.seed(123)

data_RS_sim_child_parent <-
  as_tibble(sim.poly.ideal(nvar = 50, n = 1000, cat = 4,)[["items"]]) %>%
  rename_with(~ str_c("cpi", str_pad(
    as.character(1:50), 2, side = "left", pad = "0"
  ))) %>%
  mutate(
    CPS1_raw = rowSums(.[str_c("cpi", str_pad(as.character(seq(1, 50, by = 5)), 2, side = "left", pad = "0"))]),
    CPS2_raw = rowSums(.[str_c("cpi", str_pad(as.character(seq(2, 50, by = 5)), 2, side = "left", pad = "0"))]),
    CPS3_raw = rowSums(.[str_c("cpi", str_pad(as.character(seq(3, 50, by = 5)), 2, side = "left", pad = "0"))]),
    CPS4_raw = rowSums(.[str_c("cpi", str_pad(as.character(seq(4, 50, by = 5)), 2, side = "left", pad = "0"))]),
    CPS5_raw = rowSums(.[str_c("cpi", str_pad(as.character(seq(5, 50, by = 5)), 2, side = "left", pad = "0"))]),
    CPTOT_raw = rowSums(.[str_c("cpi", str_pad(as.character(1:50), 2, side = "left", pad = "0"))]),
    across(
      contains("cpi"),
      ~ case_when(
        .x == 0 ~ "never",
        .x == 1 ~ "occasionally",
        .x == 2 ~ "frequently",
        .x == 3 ~ "always"
      )
    ),
    ID = 100001:101000,
    age = sample(c(5:12), 1000, replace = TRUE),
    age_range = case_when(age <= 8 ~ "5 to 8 yo",
                          T ~ "9 to 12 yo"),
    gender = sample(
      c("female", "male"),
      1000,
      replace = TRUE,
      prob = c(0.53, 0.47)
    ),
    educ = sample(
      c("no_HS", "HS_grad", "some_college", "BA_plus"),
      1000,
      replace = TRUE,
      prob = c(0.119, 0.263, 0.306, 0.311)
    ),
    ethnic = sample(
      c("hispanic", "asian", "black", "white", "other"),
      1000,
      replace = TRUE,
      prob = c(0.239, 0.048, 0.136, 0.521, .056)
    ),
    region = sample(
      c("northeast", "south", "midwest", "west"),
      1000,
      replace = TRUE,
      prob = c(0.166, 0.383, 0.212, 0.238)
    ),
    clin_status = sample(
      c("typ", "clin"),
      1000,
      replace = TRUE,
      prob = c(0.8, 0.2)
    )
  ) %>%
  select(ID:clin_status, CPS1_raw:CPTOT_raw, cpi01:cpi50)

write_csv(data_RS_sim_child_parent,
          here("INPUT-FILES/data-RS-sim-child-parent.csv"),
          na = "")

data_input_bfi <- bfi %>%
  drop_na() %>%
  sample_n(1000) %>%
  mutate(
    AGR_raw = rowSums(.[str_c("A", 1:5)]),
    CON_raw = rowSums(.[str_c("C", 1:5)]),
    EXT_raw = rowSums(.[str_c("E", 1:5)]),
    NEU_raw = rowSums(.[str_c("N", 1:5)]),
    OPE_raw = rowSums(.[str_c("O", 1:5)]),
    across(
      c(A1:O5),
      ~
        case_when(
          .x == 1 ~ "very_inaccurate",
          .x == 2 ~ "moderately_inaccurate",
          .x == 3 ~ "slightly_inaccurate",
          .x == 4 ~ "slightly_accurate",
          .x == 5 ~ "moderately_accurate",
          .x == 6 ~ "very_accurate",
        )
    ),
    ID = 200001:201000,
    age_range = case_when(
      age <= 18 ~ "18 yo or younger",
      between(age, 19, 24) ~ "19 to 24 yo",
      between(age, 25, 39) ~ "25 to 39 yo",
      T ~ "40 yo or older"
    ),
    gender = case_when(gender == 1 ~ "male",
                       gender == 2 ~ "female"),
    educ = case_when(
      education == 1 ~ "no_HS",
      education == 2 ~ "HS_grad",
      education == 3 ~ "some_college",
      T ~ "BA_plus"
    ),
    ethnic = sample(
      c("hispanic", "asian", "black", "white", "other"),
      1000,
      replace = TRUE,
      prob = c(0.239, 0.048, 0.136, 0.521, .056)
    ),
    region = sample(
      c("northeast", "south", "midwest", "west"),
      1000,
      replace = TRUE,
      prob = c(0.166, 0.383, 0.212, 0.238)
    ),
    clin_status = sample(
      c("typ", "clin"),
      1000,
      replace = TRUE,
      prob = c(0.8, 0.2)
    )
  ) %>%
  select(ID, age, age_range, gender:clin_status, AGR_raw:OPE_raw, A1:O5)

write_csv(data_input_bfi,
          here("INPUT-FILES/data-input-bfi.csv"),
          na = "")
```

#### Commented Snippets
Load packages for file path specification (`here`), data wrangling (`tidyverse`) and psychometric data simulation and analysis (`psych`). `base::set.seed()` ensures repeatable results from sampling and modeling functions that use a random-number generator.
```{r create-data-rsa, echo = 1:5, eval = F}
```
`sim.poly.ideal()` is a function within the `psych` package for simulating data from polytomous psychometric items. Its arguments specify the number of items `nvar`, number of cases `n`, and number of rating scale anchors `cat`. The output of `sim.poly.ideal()` is a list, within which the matrix of items and cases can be isolated with `[["items"]]` and coerced to a tibble data frame with `tibble::as_tibble()`. 
```{r create-data-rsa, echo = 8, eval = F}
```
`dplyr::rename_with()` renames columns by applying a function (here set off with the formula shortand `~`). Here we write a function to rename the fifty item-response columns with an `i01` to `i50` nomenclature. `stringr::str_pad()` pads a numerical sequence with zeros to the left of each number. Note the use of `base::as_character()` to coerce the number sequence `1:50` to a sequence of strings. This is wrapped in `stringr::str_c()` which concatenates the prefix `cpi` with the padded numbers.
```{r create-data-rsa, echo = 9:11, eval = F}
```
We use `dplyr::mutate()` to create/compute new variables. Within `mutate()`, we use `base::rowSums()` to sum across a subset of 10 items to compute the raw score  `S1_raw`. We identify the target items by their column names, including only items `i01`, `i06`, `i11`, `i16`, `i21`, `i26`, `i31`, `i36`, `i41`, and `i46` by calling `base::seq(1, 50, by = 5)`. This latter function is wrapped in the `str_c()` function described previously to concatenate the complete column names.
```{r create-data-rsa, echo = 12:13, eval = F}
```
`across()` can be used within `dplyr` verbs to delineate a subset of columns and apply a function to them. The first argument of `across()` designates the columns, using the `tidyselect` helper `contains("i")` to select only columns whose names contain `cpi`, (i.e., the item columns).

The second argument of `across()` is the function (set off by `~`) to be applied to the selected columns. Here we use `dplyr::case_when()` to recode the numerical item codes to frequency descriptors. The `.x` shorthand signifies the input cell value, which is recoded according to the predicate functions passed as arguments to `case_when()`. As an example, `.x == 0 ~ "never",` indicates that when the input cell value is 0, the value of that cell will be changed to the string "never". The formula shorthand `~` indicates the "if-then" relationship between LHS (logical predicate) and RHS (outcome).

Note that when across is used in this manner, no new columns are created; rather, existing columns are transformed. Thus, both the columns to be transformed (first argument) and the function (second argument) need to be within the parentheses of `across()`.
```{r create-data-rsa, echo = 19:27, eval = F}
```
The next snippet is an example of creating a new column `gender` and using `base::sample()` to populate it with proportionally sampled data. The arguments to `sample()` are as follows: a vector giving cell values from which to sample (`c("female", "male")`), the number of rows in the current data object (`1000`), sampling with or without replacement (`replace = TRUE`), and a vector of probability weights corresponding to each cell value (`c(0.53, 0.47)`). In this way, we compute a new variable that establishes a distribution of gender values equivalent to the U.S. census proportions for gender.
```{r create-data-rsa, echo = 32:37, eval = F}
```
We use `dplyr::select()` both to keep certain columns for the output and to specify the sequence for those columns.
```{r create-data-rsa, echo = 63, eval = F}
```
We use `readr::write_csv()` to write output to a flat file (.csv). In this instance, the arguments are:

1. `data_RS_sim_child_parent`: data object to be written
2. `here("INPUT-FILES/data-RS-sim-child-parent.csv")`: file path (specified using `here::here()`, which enables writing the file path out of the R project folder without a leading `/`)
3. `na = ""`: values that are `NA` in R will be written to the .csv as empty cells.

```{r create-data-rsa, echo = 65:67, eval = F}
```
Here we process the `bfi` data set, which has 2800 cases and is included in the `psych` package. Because we have many more cases than we need, we can clean up the data by simply dropping any row that has `NA`, which we do with `tidyr::drop_na()`. We then pare the data down to a random sample of 1000 rows, using `dplyr::sample_n()`.
```{r create-data-rsa, echo = 68:71, eval = F}
```

